---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `R`/`unihtee`

> Univariate Heterogeneous Treatment Effect Estimation

__Author:__ [Philippe Boileau](https://pboileau.ca/)

<!-- badges: start -->
<!-- badges: end -->

---

`unicate` provides tools for uncovering treatment effect modifiers in
high-dimensional data. Treatment effect modification is defined using variable
importance parameters based on risk differences or relative risks. The variable
importance parameters are model-agnostic, meaning they do not depend on specific
regression pocedures (for example, the variable importance measures of Random
Forests). Inference is performed about these variable importance measures using
nonparametric estimators. Users may use one-step or targeted maximum likelihood
estimators. Under general conditions, these estimators are double-robust, linear
and efficient.

## Installation

The package may be installed from GitHub using
[`remotes`](https://CRAN.R-project.org/package=remotes):

```{r gh-master-installation, eval=FALSE}
remotes::install_github("insightsengineering/unihtee")
```

`unihtee` is under active development. Check back often for updates.

## Usage

`unihtee()` is the only user-facing function. It can be used to perform
inference about the treatment effect modification variable importance
parameters. These parameters are defined for use with continuous, binary and
time-to-event outcomes with binary exposure variables. Variable importance
parameters are defined on both the risk difference and relative risk scales.

## Example

We simulate some observational data that contains ten confounders, of which are
two treatment effect modifiers. We then perform inference about the
risk-difference-based variable importance parameter.

```{r example}
library(unihtee)
library(MASS)
library(data.table)
library(sl3)

set.seed(510)

## create the dataset
n_obs <- 500
w <- mvrnorm(n = n_obs, mu = rep(0, 10), Sigma = diag(10))
confounder_names <- paste0("w_", seq_len(10))
colnames(w) <- confounder_names
a <- rbinom(n = n_obs, size = 1, prob = plogis(w[, 1] + w[, 2]))
y <- rnorm(n = n_obs, mean = w[, 1] + w[, 2] + a * w[, 3] - a * w[, 4])
dt <- as.data.table(cbind(w, a, y))

## targeted maximum likelihood estimates and testing procedure
unihtee(
  data = dt,
  confounders = confounder_names,
  modifiers = confounder_names,
  exposure = "a",
  outcome = "y",
  outcome_type = "continuous",
  risk_type = "risk difference",
  estimator = "tmle"
)
```

## Issues

If you encounter any bugs or have any specific feature requests, please [file
an issue](https://github.com/insightsengineering/unihtee/issues).

## Contributions

Contributions are very welcome. Interested contributors should consult our
[contribution
guidelines](https://github.com/insightsengineering/unihtee/blob/master/.github/CONTRIBUTING.md)
prior to submitting a pull request.


## License

The contents of this repository are distributed under the Apache 2.0 license.
See the
[`LICENSE.md`](https://github.com/insightsengineering/unihtee/blob/main/LICENSE.md)
and
[`LICENSE`](https://github.com/insightsengineering/unihtee/blob/main/LICENSE)
files for details.
