---
title: "Using `unihtee`"
author: "[Philippe Boileau](https://pboileau.ca)"
date: "`r Sys.Date()`"
bibliography: ../inst/REFERENCES.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{using-unihtee}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dplyr)
library(unihtee)
```

# Background

Why should you use `unihtee`? How is this different from alternative methods?

# `unihtee` in Action

Current use cases for unihtee include ... We provide details and examples below.

## Working with Continuous Outcomes

Summarize the full-data and observed-data models. Present the conditional
average treatment effect.

### Absolute Scale TEM-VIP

Present the TEM-VIP. Provide explanation. State identifiability conditions in
laymans terms.

State that there are two (three) estimators. Briefly cover their asymptotic
behaviour.

Simple example of one-step and TML estimators.

## Working with Binary Outcomes

Analogous to continuous outcome model. We can define TEM-VIPs on the absolute
scale as in the continuous outcome setting. However, often more informative to
do so on the relative scale.

### Relative Scale TEM-VIP

Present the TEM-VIP. Provide explanation. State identifiability conditions in
laymans terms. State that this parameter can be targeted in models with positive
continuous outcomes, too.

State that there are two (three) estimators. Briefly cover their asymptotic
behaviour.

Simple example of one-step and TML estimators.

## Working with Time-to-Event Outcomes

Provide models. Talk about the restricted mean survival time.

### Absolute Scale TEM-VIP

Present the TEM-VIP. Provide explanation. State identifiability conditions in
laymans terms.

State that there are two (three) estimators. Briefly cover their asymptotic
behaviour.

Simple example of one-step and TML estimators.

### Relative Scale TEM-VIP

Present the TEM-VIP. Provide explanation. State identifiability conditions in
laymans terms.

State that there are two (three) estimators. Briefly cover their asymptotic
behaviour.

Simple example of one-step and TML estimators.
