---
title: "Using `unihtee`"
author: "[Philippe Boileau](https://pboileau.ca)"
date: "`r Sys.Date()`"
bibliography: ../inst/references.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{using-unihtee}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(dplyr)
library(sl3)
library(unihtee)
```

# Background

Treatment effect modifiers (TEMs) are responsible for the disparate effects of a
treatment on a population. They are confounders --- meaning they influence both
the likelihood of receiving treatment and the outcome --- that modify the effect
of treatment on the outcome. In precision medicine, these effect modifiers
delineate patient subgroups which experience differing benefit from a given
medical intervention. In public health, they are used to determine the effect of
policy decisions on sub-populations. Identifying TEMs, if any exist, is required
for a comprehensive understanding of an intervention's effect on a population.

Traditional parametric modeling techniques, like generalized linear models
(GLMs), define TEMs as the confounders with non-zero confounder-treatment
interaction terms. In settings characterized by time-to-event outcomes, the Cox
proportional hazards model can similarly be used. Inference about TEMs is
therefore possible under stringent conditions about the data-generating process.
When these (semi)parametric methods' assumptions, like linearity of the
confounder-outcome relationship in a linear model or proportional hazards in a
Cox model, are violated, however, their results are invalid.

More flexible approaches instead focus on estimating the conditional average
treatment effect (CATE) using interpretable modeling techniques, like the LASSO
[@tibshirani1996, @tian2014, @chen2017, @zhao2018, @semenova2021,
@bahamyirou2022] or Random Forests [@breiman2001, @wager2018, @cui2022]. Again,
however, these methods require restrictive assumptions about the data-generating
process, like sparsity of treatment effect modification [@zhao2006],
approximately uncorrelated confounders [@zhao2006, @esl2009] and sample sizes
that are much larger than the number of confounders [@zhao2006], to reliably
recover TEMs.

Instead of identifying TEMs indirectly by modeling the
confounder-treatment-outcome relationship, recent work has developed frameworks
tailored to the task [@williamson2022framework, @boileau2022, @hines2022TEVIP,
@boileau2023]. Relying on TEM variable importance parameters (TEM-VIP), it is
possible to assess confounders' influence on the treatment effect in
nonparametric, algorithm-agnostic fashion that largely avoids the pitfalls of
parametric and CATE-based methodologies. These TEM-VIPs are defined within
nonparametric statistical models that may be augmented with causal
interpretations, permitting formal inference about TEMs.

The `unihtee` R package implements the recent proposals of @boileau2022 and
@boileau2023. This TEM-VIP framework relies on parameters assess the marginal
effect of each confounder on treatment effect heterogeneity, however treatment
effect is defined. Nonparametric estimators are provided in these works, and
their asymptotic properties established. In particular, these estimators are
shown to be asymptotically linear under minimal assumptions about the
data-generating process, and to recover treatment effect modifiers more readily
than competing methods. These TEM-VIPs and estimators are introduced alongside
worked examples in the following section.

# `unihtee` in Action

The `unihtee` R package performs inference about TEM-VIPs in data-generating
processes with a binary exposure variable and either continuous, binary, or
right-censored time-to-event outcomes. No restriction is placed on the
confounding variables, though ordinal confounders should be formatted as factors
using `ordered()`, and categorical confounders should be one-hot encoded.

## Working with Continuous Outcomes

Let there be $n$ independent and identically distributed (i.i.d.) random vectors
${X_i}_{i=1}^n$ such that $X_i = (W_i, A_i, Y^{(0)}, Y_i^{(1)}) \sim P_{X,0} \in
\mathcal{M}_X$. Dropping the index where possible for notational convenience for
the remainder of the tutorial, $W$ is defined as the set of $p$ confounders, $A$
as the binary treatment indicator, and $Y^{(0)}, Y_i^{(1)}$ the continuous
potential outcomes [@rubin1974] produced under control and treatment conditions,
respectively. Further, $\mathcal{M}_X$ is the full-data nonparametric model of
all possible data-generating processes, and contains the true but unknown
full-data data-generating process $P_{X,0}$.

Now, $P_{X,0}$ is generally unknown because its random vectors are unobservable:
only one potential outcome is ever measurable in practice. Still, $P_{X,0}$
allows to us to readily define causal parameters on which inference may
subsequently be performed. Examples are given in the subsections below.

We instead have access to $n$ i.i.d. random observations $O=(W,A,Y) \sim P_0 \in
\mathcal{M}$, where $W$ and $A$ are defined as in the full-data, and where $Y =
AY^{(1)} + (1-A)Y{(0)}$. $\mathcal{M}$ is the observed-data statistical model
containing all possible observed-data data-generating processes, including the
true data-generating process $P_{0}$. Note too that the elements of
$\mathcal{M}$ are fully specified by their full-data counterparts in
$\mathcal{M}_{X}$.

Having established the requisite statistical objects, let's define a
data-generating process:
```{r cont-dgp}
cont_outcome_dgp <- function(n_obs) {

  # confounders
  w_1 <- rnorm(n = n_obs)
  w_2 <- rnorm(n = n_obs)
  w_3 <- rnorm(n = n_obs)
  w_4 <- rnorm(n = n_obs)
  w_5 <- rnorm(n = n_obs)

  # treatment
  prop_score <- plogis(w_1 + w_2)
  a <- sapply(prop_score, function(x) rbinom(1, 1, prob = x))

  # potential outcomes
  y_1 <- rnorm(n = n_obs, mean = w_1 + 2 * w_2 + 2 * w_3 + 0.5 * a, sd = 0.1)
  y_0 <- rnorm(n = n_obs, mean = w_1 + w_2, sd = 0.1)

  # outcome
  y <- a * y_1 + (1 - a) * y_0

  # assemble the observations in a tibble
  tibble(
    w_1 = w_1,
    w_2 = w_2,
    w_3 = w_3,
    w_4 = w_4,
    w_5 = w_5,
    a = a,
    y = y
  )
}
```

It's clear from the definition of the potential outcomes that `w_2` and `w_3`
are treatment effect modifiers: they interact with the treatment indicator `a`.
A TEM-VIP for quantifying the strength of the treatment effect modification is
outlined next.

### Absolute TEM-VIP

Indexing $W$ by $j=1,\ldots,p$, and assuming without
loss of generality that
$\mathbb{E}_{P_{X,0}}[W_j] = 0, \mathbb{E}_{P_{X,0}}[W_j^2] > 0$, we can define
our first TEM-VIP for the $j^\text{th}$ confounder as follows:

$$
\Psi^F_j(P_{X,0}) =
\frac{\mathbb{E}_{P_{X,0}}[(Y^{(1)}-Y^{(0)})W_j]}{\mathbb{E}_{P_{X,0}}[W_j^2]}
=\frac{\mathbb{E}_{P_{X,0}}[(\mathbb{E}_{P_{X,0}}[Y^{(1)}|W]-
  \mathbb{E}_{P_{X,0}}[Y^{(0)}|W])W_j]}
  {\mathbb{E}_{P_{X,0}}[W_j^2]} \;.
$$

This call this parameter the absolute TEM-VIP. Assuming the expectation of
$\mathbb{E}_{P_{X,0}}[Y^{(1)}|W]- \mathbb{E}_{P_{X,0}}[Y^{(0)}|W]$ conditional
on $W_j$ is linear in $W_j$, $\Psi^F_j(P_{X,0})$ is the simple linear regression
coefficient produced by regressing the difference in expected potential outcomes
against confounder $W_j$. Even when this relationship is nonlinear, as is almost
surely the case in most applications, $\Psi^F_j(P_{X,0})$ corresponds to the
correlation between the difference in potential outcomes and the $j^\text{th}$
confounder, re-normalized to be on the same scale as the potential outcomes.

Of course, as stated earlier, we don't generally have access to the $X$s
required to perform inference about the above parameter. Luckily, under the
identification conditions outlined in @boileau2022 and @boileau2023 --- namely,
that there is no unmeasured confounding and no positivity violations --- we can
perform inference about the equivalent observed-data parameter:

$$
\Psi_j(P_0)
=\frac{\mathbb{E}_{P_0}[(\mathbb{E}_{P_{0}}[Y|A=1,W]-
  \mathbb{E}_{P_0}[Y|A=0,W])W_j]}{\mathbb{E}_{P_0}[W_j^2]} 
= \Psi^F_j(P_{X,0})\;.
$$

@boileau2022 and @boileau2023 derive two nonparametric estimators of this
parameter: the one-step and targeted maximum likelihood (TML) estimators. Both
require the estimation of two nuisance parameters: the propensity score and the
expected outcome conditioned on the treatment and confounders. These estimators
are double robust: only one nuisance parameter is required to be consistently
estimated to ensure that the one-step and TML estimators are consistent. If both
nuisance parameters converge to their true values at a fast enough rate, then
the one-step and TML estimators are asymptotically normal. This permits
hypothesis testing about $\Psi_j(P_0)$ using Wald-type confidence intervals.
Further details on these estimators and their asymptotic properties are provided
in @boileau2022 and @boileau2023.

Having defined the parameter and briefly discussed the estimators, we apply
`unicate()` to recover the treatment effect modifiers, as defined by the
absolute TEM-VIP. Both the one-step and TML estimators are showcased. The LASSO
regression of @tibshirani1996 is used to estimate the nuisance parameters.

```{r abs-tem-vip}
set.seed(514)

# simulate a random sample
sample_df <- cont_outcome_dgp(n_obs = 500)

# one-step estimator
unihtee(
  data = sample_df,
  confounders = c("w_1", "w_2", "w_3", "w_4", "w_5"),
  modifiers =  c("w_1", "w_2", "w_3", "w_4", "w_5"),
  exposure = "a",
  outcome = "y",
  outcome_type = "continuous",
  scale = "absolute",
  estimator = "onestep",
  cond_outcome_estimator = sl3::Lrnr_glmnet$new(
    family = "gaussian",
    formula = "~ a * w_1 + a * w_2 + a * w_3 + a * w_4 + a * w_5"
  ),
  prop_score_estimator = sl3::Lrnr_glmnet$new(family = "binomial")
)

# TML estimator
unihtee(
  data = sample_df,
  confounders = c("w_1", "w_2", "w_3", "w_4", "w_5"),
  modifiers =  c("w_1", "w_2", "w_3", "w_4", "w_5"),
  exposure = "a",
  outcome = "y",
  outcome_type = "continuous",
  scale = "absolute",
  estimator = "tmle",
  cond_outcome_estimator = sl3::Lrnr_glmnet$new(
    family = "gaussian",
    formula = "~ a * w_1 + a * w_2 + a * w_3 + a * w_4 + a * w_5"
  ),
  prop_score_estimator = sl3::Lrnr_glmnet$new(family = "binomial")
)
```

`unicate()` outputs a table of results summarizing the TEM-VIP inference
procedure for each potential treatment effect modifier. This table ordered by
the nominal p-value of the inference procedure. Both the one-step and TML
estimators correctly identify the TEM-VIP of `w_2` and `w_3` as being non-zero
based on the false discovery rate adjusted p-values. Each estimator's point
estimates are also close to their true parameter values.

## Working with Binary Outcomes

Analogous to continuous outcome model. We can define TEM-VIPs on the absolute
scale as in the continuous outcome setting. However, often more informative to
do so on the relative scale.

### Relative TEM-VIP

Present the TEM-VIP. Provide explanation. State identifiability conditions in
laymans terms. State that this parameter can be targeted in models with positive
continuous outcomes, too.

State that there are two (three) estimators. Briefly cover their asymptotic
behaviour.

Simple example of one-step and TML estimators.

## Working with Time-to-Event Outcomes

Provide models. Talk about the restricted mean survival time.

### Absolute TEM-VIP

Present the TEM-VIP. Provide explanation. State identifiability conditions in
laymans terms.

State that there are two (three) estimators. Briefly cover their asymptotic
behaviour.

Simple example of one-step and TML estimators.

### Relative TEM-VIP

Present the TEM-VIP. Provide explanation. State identifiability conditions in
laymans terms.

State that there are two (three) estimators. Briefly cover their asymptotic
behaviour.

Simple example of one-step and TML estimators.

# Session information

```{r session-information}
sessionInfo()
```
