---
title: "Using `unihtee`"
author: "[Philippe Boileau](https://pboileau.ca)"
date: "`r Sys.Date()`"
bibliography: ../inst/references.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{using-unihtee}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(dplyr)
library(sl3)
library(unihtee)
```

# Background

Treatment effect modifiers (TEMs) are responsible for the disparate effects of a
treatment on a population. They are confounders --- meaning they influence both
the likelihood of receiving treatment and the outcome --- that modify the effect
of treatment on the outcome. In precision medicine, these effect modifiers
delineate patient subgroups which experience differing benefit from a given
medical intervention. In public health, they are used to determine the effect of
policy decisions on sub-populations. Identifying TEMs, if any exist, is required
for a comprehensive understanding of an intervention's effect on a population.

Traditional parametric modeling techniques, like generalized linear models
(GLMs), define TEMs as the confounders with non-zero confounder-treatment
interaction terms. In settings characterized by time-to-event outcomes, the Cox
proportional hazards model can similarly be used. Inference about TEMs is
therefore possible under stringent conditions about the data-generating process.
When these (semi)parametric methods' assumptions, like linearity of the
confounder-outcome relationship in a linear model or proportional hazards in a
Cox model, are violated, however, their results are invalid.

More flexible approaches instead focus on estimating the conditional average
treatment effect (CATE) using interpretable modeling techniques, like the LASSO
[@tibshirani1996; @tian2014; @chen2017; @zhao2018; @semenova2021;
@bahamyirou2022] or Random Forests [@breiman2001; @wager2018; @cui2022]. Again,
however, these methods require restrictive assumptions about the data-generating
process, like sparsity of treatment effect modification [@zhao2006],
approximately uncorrelated confounders [@zhao2006; @esl2009] and sample sizes
that are much larger than the number of confounders [@zhao2006], to reliably
recover TEMs.

Instead of identifying TEMs indirectly by modeling the
confounder-treatment-outcome relationship, recent work has developed frameworks
tailored to the task [@williamson2022framework; @boileau2022; @hines2022TEVIP;
@boileau2023]. Relying on TEM variable importance parameters (TEM-VIP), it is
possible to assess confounders' influence on the treatment effect in
nonparametric, algorithm-agnostic fashion that largely avoids the pitfalls of
parametric and CATE-based methodologies. These TEM-VIPs are defined within
nonparametric statistical models that may be augmented with causal
interpretations, permitting formal inference about TEMs.

The `unihtee` R package implements the recent proposals of @boileau2022 and
@boileau2023. This TEM-VIP framework relies on parameters assess the marginal
effect of each confounder on treatment effect heterogeneity, however treatment
effect is defined. Nonparametric estimators are provided in these works, and
their asymptotic properties established. In particular, these estimators are
shown to be asymptotically linear under minimal assumptions about the
data-generating process, and to recover treatment effect modifiers more readily
than competing methods. These TEM-VIPs and estimators are introduced alongside
worked examples in the following section.

# `unihtee` in Action

The `unihtee` R package performs inference about TEM-VIPs in data-generating
processes with a binary exposure variable and either continuous, binary, or
right-censored time-to-event outcomes. No restriction is placed on the
confounding variables, though ordinal confounders should be formatted as factors
using `ordered()`, and categorical confounders should be one-hot encoded.

## Working with Continuous Outcomes

Let there be $n$ independent and identically distributed (i.i.d.) random vectors
${X_i}_{i=1}^n$ such that $X_i = (W_i, A_i, Y^{(0)}, Y_i^{(1)}) \sim P_{X,0} \in
\mathcal{M}_X$. Dropping the index where possible for notational convenience for
the remainder of the tutorial, $W$ is defined as the set of $p$ confounders, $A$
as the binary treatment indicator, and $Y^{(0)}, Y_i^{(1)}$ the continuous
potential outcomes [@rubin1974] produced under control and treatment conditions,
respectively. Further, $\mathcal{M}_X$ is the full-data nonparametric model of
all possible data-generating processes, and contains the true but unknown
full-data data-generating process $P_{X,0}$.

Now, $P_{X,0}$ is generally unknown because its random vectors are unobservable:
only one potential outcome is ever measurable in practice. Still, $P_{X,0}$
allows to us to readily define causal parameters on which inference may
subsequently be performed. Examples are given in the subsections below.

We instead have access to $n$ i.i.d. random observations $O=(W,A,Y) \sim P_0 \in
\mathcal{M}$, where $W$ and $A$ are defined as in the full-data, and where $Y =
AY^{(1)} + (1-A)Y{(0)}$. $\mathcal{M}$ is the observed-data statistical model
containing all possible observed-data data-generating processes, including the
true data-generating process $P_{0}$. Note too that the elements of
$\mathcal{M}$ are fully specified by their full-data counterparts in
$\mathcal{M}_{X}$.

Having established the requisite statistical objects, let's define a
data-generating process:
```{r cont-dgp}
cont_outcome_dgp <- function(n_obs) {

  # confounders
  w_1 <- rnorm(n = n_obs)
  w_2 <- rnorm(n = n_obs)
  w_3 <- rnorm(n = n_obs)
  w_4 <- rnorm(n = n_obs)
  w_5 <- rnorm(n = n_obs)

  # treatment
  prop_score <- plogis(w_1 + w_2)
  a <- rbinom(n_obs, 1, prob = prop_score)

  # potential outcomes
  y_1 <- rnorm(n = n_obs, mean = w_1 + 2 * w_2 + 2 * w_3 + 0.5 * a, sd = 0.1)
  y_0 <- rnorm(n = n_obs, mean = w_1 + w_2, sd = 0.1)

  # outcome
  y <- a * y_1 + (1 - a) * y_0

  # assemble the observations in a tibble
  tibble(
    w_1 = w_1,
    w_2 = w_2,
    w_3 = w_3,
    w_4 = w_4,
    w_5 = w_5,
    a = a,
    y = y
  )
}
```

It's clear from the definition of the potential outcomes that `w_2` and `w_3`
are treatment effect modifiers: they interact with the treatment indicator `a`.
A TEM-VIP for quantifying the strength of the treatment effect modification is
outlined next.

### Additive TEM-VIP

Indexing $W$ by $j=1,\ldots,p$, and assuming without
loss of generality that
$\mathbb{E}_{P_{X,0}}[W_j] = 0, \mathbb{E}_{P_{X,0}}[W_j^2] > 0$, we can define
our first TEM-VIP for the $j^\text{th}$ confounder as follows:

$$
\Psi^F_j(P_{X,0}) =
\frac{\mathbb{E}_{P_{X,0}}[(Y^{(1)}-Y^{(0)})W_j]}{\mathbb{E}_{P_{X,0}}[W_j^2]}
=\frac{\mathbb{E}_{P_{X,0}}[(\mathbb{E}_{P_{X,0}}[Y^{(1)}|W]-
  \mathbb{E}_{P_{X,0}}[Y^{(0)}|W])W_j]}
  {\mathbb{E}_{P_{X,0}}[W_j^2]} \;.
$$

This parameter is called the additive TEM-VIP. Assuming the expectation of
$\mathbb{E}_{P_{X,0}}[Y^{(1)}|W]- \mathbb{E}_{P_{X,0}}[Y^{(0)}|W]$ conditional
on $W_j$ is linear in $W_j$, $\Psi^F_j(P_{X,0})$ is the simple linear regression
coefficient produced by regressing the difference in expected potential outcomes
against confounder $W_j$. Even when this relationship is nonlinear, as is almost
surely the case in most applications, $\Psi^F_j(P_{X,0})$ corresponds to the
correlation between the difference in potential outcomes and the $j^\text{th}$
confounder, re-normalized to be on the same scale as the potential outcomes.

Of course, as stated earlier, we don't generally have access to the full-data
random vectors required to perform inference about the above parameter. Luckily,
under the identification conditions outlined in @boileau2022 and @boileau2023
--- namely, that there is no unmeasured confounding and no positivity violations
--- we can perform inference about the equivalent observed-data parameter:

$$
\Psi_j(P_0)
=\frac{\mathbb{E}_{P_0}[(\mathbb{E}_{P_{0}}[Y|A=1,W]-
  \mathbb{E}_{P_0}[Y|A=0,W])W_j]}{\mathbb{E}_{P_0}[W_j^2]} 
= \Psi^F_j(P_{X,0})\;.
$$

@boileau2022 and @boileau2023 derive two nonparametric estimators of this
parameter: the one-step and targeted maximum likelihood (TML) estimators. Both
require the estimation of two nuisance parameters: the propensity score and the
expected outcome conditioned on the treatment and confounders. These estimators
are double robust: only one nuisance parameter is required to be consistently
estimated to ensure that the one-step and TML estimators are consistent. If both
nuisance parameters converge to their true values at a fast enough rate, then
the one-step and TML estimators are asymptotically normal. This permits
hypothesis testing about $\Psi_j(P_0)$ using Wald-type confidence intervals.
Further details on these estimators and their asymptotic properties are provided
in @boileau2022 and @boileau2023.

Having defined the parameter and briefly discussed the estimators, we apply
`unicate()` to recover the treatment effect modifiers, as defined by the
additive TEM-VIP. Both the one-step and TML estimators are showcased. The LASSO
regression of @tibshirani1996 is used to estimate the nuisance parameters.

```{r additive-tem-vip}
set.seed(514)

# simulate a random sample
sample_df <- cont_outcome_dgp(n_obs = 500)

# one-step estimator
unihtee(
  data = sample_df,
  confounders = c("w_1", "w_2", "w_3", "w_4", "w_5"),
  modifiers =  c("w_1", "w_2", "w_3", "w_4", "w_5"),
  exposure = "a",
  outcome = "y",
  outcome_type = "continuous",
  scale = "absolute",
  estimator = "onestep",
  cond_outcome_estimator = sl3::Lrnr_glmnet$new(
    family = "gaussian",
    formula = "~ a * w_1 + a * w_2 + a * w_3 + a * w_4 + a * w_5"
  ),
  prop_score_estimator = sl3::Lrnr_glmnet$new(family = "binomial")
)

# TML estimator
unihtee(
  data = sample_df,
  confounders = c("w_1", "w_2", "w_3", "w_4", "w_5"),
  modifiers =  c("w_1", "w_2", "w_3", "w_4", "w_5"),
  exposure = "a",
  outcome = "y",
  outcome_type = "continuous",
  scale = "absolute",
  estimator = "tmle",
  cond_outcome_estimator = sl3::Lrnr_glmnet$new(
    family = "gaussian",
    formula = "~ a * w_1 + a * w_2 + a * w_3 + a * w_4 + a * w_5"
  ),
  prop_score_estimator = sl3::Lrnr_glmnet$new(family = "binomial")
)
```

`unicate()` outputs a table of results summarizing the TEM-VIP inference
procedure for each potential treatment effect modifier. This table ordered by
the nominal p-value of the inference procedure. Both the one-step and TML
estimators correctly identify the TEM-VIP of `w_2` and `w_3` as being non-zero
based on the false discovery rate adjusted p-values. Each estimator's point
estimates are also close to their true parameter values.


## Working with Binary Outcomes

In the binary outcome setting, the full-data and observed-data models and
data-generating processes are identical to those of the continuous outcome
setting, save that the outcome is binary. That means the additive TEM-VIP can be
used as is with binary-outcome data --- the only practical difference is the
choice of estimator used for the expected outcome conditional on the treatment
and confounders. With that said, a relative TEM-VIP is generally more sensitive
and informative in these scenarios. We propose such a parameter in the following
section.

Before that, however, we define a data-generating process with two treatment
effect modifiers:

```{r binary-dgp}
bin_outcome_dgp <- function(n_obs) {

  # confounders
  w_1 <- rpois(n = n_obs, lambda = 3)
  w_2 <- rnorm(n = n_obs)
  w_3 <- rbinom(n = n_obs, size = 1, prob = 0.5)

  # treatment
  prop_score <- plogis(-0.5 + 0.25 * w_1 + w_2 + 0.5 * w_3)
  a <- rbinom(n_obs, 1, prob = prop_score)

  # potential outcomes
  y_prob_1 <- plogis(-1 - 3 * w_3)
  y_prob_0 <- plogis(1 + 2 * w_2)
  y_1 <- rbinom(n_obs, 1, prob = y_prob_1)
  y_0 <- rbinom(n_obs, 1, prob = y_prob_0)

  # outcome
  y <- a * y_1 + (1 - a) * y_0

  # assemble the observations in a tibble
  tibble(
    w_1 = w_1,
    w_2 = w_2,
    w_3 = w_3,
    a = a,
    y = y
  )
}
```

Here, `w_2` and `w_3` are a treatment effect modifier. They interact with the
treatment variable, `a`.


### Relative TEM-VIP

Again assuming that the confounders are centered at zero with non-zero variance,
the relative TEM-VIP for the $j^\text{th}$ confounder is
$$
\Gamma^F_j(P_{X,0})
=\frac{\mathbb{E}_{P_{X,0}}[(\log\mathbb{E}_{P_{X,0}}[Y^{(1)}|W]-
  \log\mathbb{E}_{P_{X,0}}[Y^{(0)}|W])W_j]}
  {\mathbb{E}_{P_{X,0}}[W_j^2]} \;.
$$

Assuming that the expectation of the log ratio of the expected conditional
potential outcomes conditional on $W_j$ is linear in $W_j$,
$\Gamma^F_j(P_{X,0})$ is the simple linear regression coefficient obtained by
regressing the log ratio of the expected conditional potential outcomes on
$W_j$. As with $\Psi^F_j(P_{X,0})$, this parameter can be interpreted as
standardized correlation coefficient of the log ratio of expected potential
outcomes and the $j^\text{th}$ confounder.

We highlight that this TEM-VIP isn't restricted to data-generating processes
with binary outcomes; it may also be used when the outcome is non-negative,
continuous or discrete random variable.

Again, much like $\Psi^F_j(P_{X,0})$, $\Gamma^F_j(P_{X,0})$ isn't generally ever
observed since it is a parameter of the full-data model. Under the conditions of
no unmeasured confounding and in the absence of positivity violations, however,
an equivalent parameter can be estimated from the observed-data:

$$
\Gamma_j(P_0)
=\frac{\mathbb{E}_{P_0}[(\log\mathbb{E}_{P_{0}}[Y|A=1,W]-
  \log\mathbb{E}_{P_0}[Y|A=0,W])W_j]}{\mathbb{E}_{P_0}[W_j^2]} 
= \Gamma^F_j(P_{X,0})\;.
$$

Nonparametric one-step and TML estimators of this parameter are derived by
@boileau2023 and implemented in the `unicate` package. Just like $\Psi_j(P_0)$,
inference about this parameter depends on the estimation of the propensity score
and the expected outcome conditional on treatment indicator and confounders.
These estimators are consistent if both nuisance parameters are consistently
estimated, and are asymptotically normal if both converge to their true values
at rate of $n^{-1/4}$. The latter permits hypothesis testing based on the
Gaussian. Additional details are provided by @boileau2023.

We now apply these estimators to a random sample generated by
`bin_outcome_dgp()`. The nuisance parameters are estimated using Random Forests
implemented in `ranger` package [@breiman2001; @ranger]. Note too that these
estimators are fit using cross-fitting [@zheng2011cross; @cherno2017].

```{r relative-tem-vip}
# simulate a random sample
sample_df <- bin_outcome_dgp(n_obs = 500)

# one-step estimator
unihtee(
  data = sample_df,
  confounders = c("w_1", "w_2", "w_3"),
  modifiers =  c("w_1", "w_2", "w_3"),
  exposure = "a",
  outcome = "y",
  outcome_type = "binary",
  scale = "relative",
  estimator = "onestep",
  cond_outcome_estimator = sl3::Lrnr_ranger$new(),
  prop_score_estimator = sl3::Lrnr_ranger$new(),
  cross_fit = TRUE
)

# TML estimator
unihtee(
  data = sample_df,
  confounders = c("w_1", "w_2", "w_3"),
  modifiers =  c("w_1", "w_2", "w_3"),
  exposure = "a",
  outcome = "y",
  outcome_type = "binary",
  scale = "relative",
  estimator = "tmle",
  cond_outcome_estimator = sl3::Lrnr_ranger$new(),
  prop_score_estimator = sl3::Lrnr_ranger$new(),
  cross_fit = TRUE
)
```

Both estimators correctly identify `w_2` and `w_3` as treatment effect
modifiers, as defined by the relative TEM-VIP. Note that the TML estimators'
estimates of `w_2` and `w_3` is noticeably smaller than that produced by the
one-step estimator; @boileau2023 have shown in a simulation study of this
estimator that the TML estimator exhibits larger finite-sample bias than the
one-step estimator. We therefore recommend using the one-step estimator in
smaller samples.

## Working with Time-to-Event Outcomes

`unihtee` also provides functionality for estimating TEM-VIPs for data with
right-censored time-to-event outcomes. We detail the assumed full-data and
observed-data models here, then introduce an additive and a relative TEM-VIP in
subsequent sections.

Consider $n$ i.i.d. random vectors $X = (W, A, C^{(0)}, C^{(1)}, T^{(0)},
T^{(1)}) \sim P_{X,0} \in \mathcal{M}_X$. $W, A, P_{X,0},$ and $\mathcal{M}_X$
are defined as in the continuous and binary outcome models. $C^{(0)}$ and
$C^{(1)}$ correspond to the censoring times that would be observed under the
control and treatment conditions, respectively. Similarly, $T^{(0)}$ and
$T^{(1)}$ are the event times that would be observed under either treatment
assignment.

Again, $P_{X,0}$ is generally unobservable. We instead have access to random
observations $O = (W, A, \Delta, T) \sim P_0 \in \mathcal{M}$, where $\Delta$ is
a censoring indicator and $T = A(\Delta C^{(1)} + (1-\Delta)T^{(1)}) +
(1-A)(\Delta C^{(0)} + (1-\Delta) T^{(0)})$. $W, A, P_0,$ and $\mathcal{M}$ are
defined as in the previous observed-data models.

Now, it's worth introducing the full-data and observed-data conditional survival
functions since both TEM-VIPs introduced in this section rely on them. The
full-data survival function is defined as $S_{P_{X,0}}(t|A,W) =
\mathbb{P}_{P_{X,0}}[T^{(A)} > t | W]$. This parameter's observed-data
counterpart is defined analogously as $S_0(t|A,W) = P_{P_0}[T > t |A, W]$.

Finally, we define a time-to-event data-generating process with nine equidistant
time points and three potential confounders, `w_1`, `w_2`, and `w_3`. Of these
confounders, only `w_1` is a treatment effect modifier. Note that `unihtee()`
requires a "wide" data format when working time-to-event outcome data. That is,
each observation is represented by a single row in the data. An example is
provided following this code chunk.

```{r tte-dgp}
# define hazard functions
cond_cens_hazard <- function(time, a, w_1, w_2, w_3) {
  (time < 9) / (1 + exp(4 + 0.2 * (w_1 + w_2) - a))
}
cond_surv_hazard <- function(time, a, w_1, w_2, w_3) {
  (time < 9) / (1 + exp(2 + 3 * a * w_1)) + (time == 9)
}

# failure time simulator
failure_time_sim <- function(n_obs, a, w_1, w_2, w_3) {
  sapply(
    seq_len(n_obs),
    function(obs) {
      failure_time <- NA
      for (t in 1:9) {
        prob <- cond_surv_hazard(t, a, w_1[obs], w_2[obs], w_3[obs])
        status <- rbinom(1, 1, prob)
        if (status == 1) {
          failure_time <- t
          break
        }
      }
      return(failure_time)
    }
  )
}

# censoring time simulator
censor_time_sim <- function(n_obs, a, w_1, w_2, w_3) {
  sapply(
    seq_len(n_obs),
    function(obs) {
      censor_time <- NA
      for (t in 1:9) {
        prob <- cond_cens_hazard(t, a, w_1[obs], w_2[obs], w_3[obs])
        status <- rbinom(1, 1, prob)
        if (status == 1) {
          censor_time <- t
          break
        }
      }
      if (is.na(censor_time)) censor_time <- 10
      return(censor_time)
    }
  )
}

# time-to-event outcome data-generating process
tte_outcome_dgp <- function(n_obs) {

  # confounders
  w_1 <- rnorm(n = n_obs)
  w_2 <- rnorm(n = n_obs)
  w_3 <- rnorm(n = n_obs)

  # treatment
  prop_score <- plogis(-1 - w_1 + w_2 - w_3)
  a <- rbinom(n_obs, 1, prob = prop_score)

  # generate the failure events for t = 1 to 9
  failure_time_1 <- failure_time_sim(n_obs, 1, w_1, w_2, w_3)
  failure_time_0 <- failure_time_sim(n_obs, 0, w_1, w_2, w_3)

  # generate the censoring events for t = 1 to 9
  censor_time_1 <- censor_time_sim(n_obs, 1, w_1, w_2, w_3)
  censor_time_0 <- censor_time_sim(n_obs, 0, w_1, w_2, w_3)

  # compile the failure and censoring times
  failure_time <- sapply(
    seq_len(n_obs),
    function(obs) {
      if (a[obs] == 1) failure_time_1[obs] else failure_time_0[obs]
    }
  )
  censor_time <- sapply(
    seq_len(n_obs),
    function(obs) {
      if (a[obs] == 1) censor_time_1[obs] else censor_time_0[obs]
    }
  )
  
  # determine the observed time-to-event and censoring indicator
  time <- sapply(
    seq_len(n_obs),
    function(obs) {
      if (censor_time[obs] < failure_time[obs]) {
        censor_time[obs]
      } else {
        failure_time[obs]
      }
    }
  )
  censoring_indicator <- sapply(
    seq_len(n_obs),
    function(obs) if (time[obs] == censor_time[obs]) 1 else 0
  )
  
  # assemble the tibble
  tibble(
    w_1 = w_1,
    w_2 = w_2,
    w_3 = w_3,
    a = a,
    time = time,
    censoring_indicator = censoring_indicator
  )
}
```

```{r tte-data-example}
tte_outcome_dgp(10)
```

### Additive TEM-VIP

@boileau2023 Propose an additive TEM-VIP based on the restricted mean survival
time at some pre-specified time $t$. Again assuming without loss of generality
that the $\mathbb{E}_{P_{X,0}}[W_j] = 0$ and that the confounders have variance
bounded away from zero, this TEM-VIP is defined as
$$
\Psi_j^F(P_{X,0};t) = \frac{\mathbb{E}_{P_{X,0}}\left[W_j
  \int_0^t \{ S_{P_{X,0}}(u|1, W) - S_{P_{X,0}}(u|0, W) \} du \right]}
  {\mathbb{E}_{P_{X,0}}[W_j^2]} \; .
$$

"$\Psi$" is re-used to highlight that this is an additive TEM-VIP. As in the
continuous outcome scenario, this parameter captures the correlation between the
conditional difference in the restricted mean survival time and the
$j^\text{th}$ confounder, re-normalized to be on the same scale as the outcome.
Put another way, this parameter identifies confounders driving the largest
difference in truncated survival times across treatment conditions.

As before, this full-data parameter is equal to the following observed-data
parameter under the assumptions of (1) no unmeasured confounding, (2) censoring
mechanism positivity, and (3) treatment assignment positivity:
$$
\Psi_j(P_0;t) = \frac{\mathbb{E}_{P_0}\left[W_j
  \int_0^t \{ S_0(u|1, W) - S_0(u|0, W) \} du \right]}
  {\mathbb{E}_{P_0}[W_j^2]} \; .
$$

As in the continuous and binary outcome scenarios, @boileau2023 derived two
estimators of this additive effect parameter: a one-step estimator and a TML
estimator. Both rely on the accurate estimation of three nuisance parameters:
the conditional event hazard function, the conditional censoring hazard
function, and the propensity score. If either the conditional event hazard
function or the conditional censoring hazard function and the propensity score
are consistently estimated, then these nonparametric estimators are consistent.
Further, if all of these nuisance parameters are converge to their true values
at a rate of $n^{-1/4}$, then these estimators are asymptotically normally
distributed. This permits Gaussian-based hypothesis testing through the use of
Wald-type confidence intervals. See @boileau2023 for more details.

```{r tte-additive-tem-vip}
# simulate a random sample
sample_df <- tte_outcome_dgp(n_obs = 250)

# one-step estimator
unihtee(
  data = sample_df,
  confounders = c("w_1", "w_2", "w_3"),
  modifiers = c("w_1", "w_2", "w_3"),
  exposure = "a",
  outcome = "time",
  censoring = "censoring_indicator",
  time_cutoff = 8,
  outcome_type = "time-to-event",
  scale = "absolute",
  estimator = "onestep",
  prop_score_estimator = sl3::Lrnr_glmnet$new(),
  failure_hazard_estimator = sl3::Lrnr_xgboost$new(),
  censoring_hazard_estimator = sl3::Lrnr_glmnet$new()
)

# TML estimator
unihtee(
  data = sample_df,
  confounders = c("w_1", "w_2", "w_3"),
  modifiers = c("w_1", "w_2", "w_3"),
  exposure = "a",
  outcome = "time",
  censoring = "censoring_indicator",
  time_cutoff = 8,
  outcome_type = "time-to-event",
  scale = "absolute",
  estimator = "tmle",
  prop_score_estimator = sl3::Lrnr_glmnet$new(),
  failure_hazard_estimator = sl3::Lrnr_xgboost$new(),
  censoring_hazard_estimator = sl3::Lrnr_glmnet$new()
)
```

### Relative TEM-VIP

Present the TEM-VIP. Provide explanation. State identifiability conditions in
laymans terms.

State that there are two (three) estimators. Briefly cover their asymptotic
behaviour.

```{r tte-relative-tem-vip}
# simulate a random sample
sample_df <- tte_outcome_dgp(n_obs = 1000)

# one-step estimator
unihtee(
  data = sample_df,
  confounders = c("w_1", "w_2", "w_3"),
  modifiers = c("w_1", "w_2", "w_3"),
  exposure = "a",
  outcome = "time",
  censoring = "censoring_indicator",
  time_cutoff = 5,
  outcome_type = "time-to-event",
  scale = "relative",
  estimator = "onestep",
  prop_score_estimator = sl3::Lrnr_glmnet$new(),
  failure_hazard_estimator = sl3::Lrnr_xgboost$new(),
  censoring_hazard_estimator = sl3::Lrnr_glmnet$new(),
  cross_fit = TRUE
)

# TML estimator
unihtee(
  data = sample_df,
  confounders = c("w_1", "w_2", "w_3"),
  modifiers = c("w_1", "w_2", "w_3"),
  exposure = "a",
  outcome = "time",
  censoring = "censoring_indicator",
  time_cutoff = 5,
  outcome_type = "time-to-event",
  scale = "relative",
  estimator = "tmle",
  prop_score_estimator = sl3::Lrnr_glmnet$new(),
  failure_hazard_estimator = sl3::Lrnr_xgboost$new(),
  censoring_hazard_estimator = sl3::Lrnr_glmnet$new(),
  cross_fit = TRUE
)
```

# Additional Functionality

## Working with Randomized Experiment Data

## Cross-Fitting

## `sl3` Integration: Using Super Learners

# Session information

```{r session-information}
sessionInfo()
```
